---
title: "NFL QB Performance - Summary"
author: "Oren Shapira"
date: "July 7, 2019"
always_allow_html: yes
output: 
  html_document:
    keep_md: yes
    toc: true
    theme: united
  github_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```


# **Summary of Project**

## Overview
With the increasing ubiquity of sports gambling and Fantasy leagues of various formats, NFL Sundays can become a stessful time for high-stakes betting. There are many publicly available tools that can guide people's betting decisions (e.g. Vegas odds, Fantasy League Point Projections), but these are often black-box algorithms that may provide contradictory information. For example, why might Yahoo Fantasy project more points for a given player than in an ESPN Fantasy league with the same scoring settings? The first step to answering this is to identify all elements that may impact a player's performance, and determine which ones have the most significant impact. My goal is to take a bite-size approach to this task. 

For this analysis, I used various machine learning algorithms to determine how much weather and and opposing team defense impact a Quarterback's performance for a given game. I defined "performance" as how many more/less passing yards a QB threw in a game compared to their season average. The specific attributes I focused on for weather were temperature, humidity, and wind (mph). For opponent defense, I focused on the relative strength of a team's passing defense.


## The Data:

This analysis required scrubbing and merging various datasets.

1. Weather_2013_1231.csv: (Weather dataset)
  Summary: This dataset contains weather information for every NFL game between 1960-2013. It has 11 attributes, and 11,192 tuples.
  Source: http://www.nflsavant.com/about.php
  Data snapshot:

```{r}
library(readxl)
weather_raw <- read.csv("C:\\Users\\oshapira\\Desktop\\Analytics\\NFL_Quarterback_Performance\\Raw_Data\\weather_20131231.csv")

library(knitr)
kable(head(weather_raw,3))

```


2. Game_Logs_Quarterback.csv (QB Stats for every game)
  Summary: This dataset contains passing statistics for every Quarterback, for NFL game between 2000 - 2016. This includes total passing yards, pass attempts, Touchdowns, Interceptions, etc. This will be joined with the weather dataset above on overlapping years. This dataset has 29 attributes and 40,247 rows. 
  Source: https://www.kaggle.com/kendallgillies/nflstatistics
  Data snapshot:
  
```{r}
game_logs_raw <- read.csv("C:\\Users\\oshapira\\Desktop\\Analytics\\NFL_Quarterback_Performance\\Raw_Data\\game_logs_Quarterback.csv")
kable(head(game_logs_raw,3))
```


3. Career_Stats_Passing.csv (QB stats for season averages)
  Summary: This dataset contains every NFL player's overall passing stats for the entire season, and spans between 1924-2016. It will be used primarily to normalize across Quarterbacks that have different averages of passing yards per game. There are 23 attributes and 8,525 rows.
  Source: https://www.kaggle.com/kendallgillies/nflstatistics
  Data snapshot:
  
```{r}
career <- read.csv("C:\\Users\\oshapira\\Desktop\\Analytics\\NFL_Quarterback_Performance\\Raw_Data\\Career_Stats_Passing.csv")
kable(head(career,3))
```
  

4. Sportsref_download_ [2000-2013].xls (Team Defense Stats)
  Summary: This is a series of 13 datasets each NFL season between 2000 - 2013. It contains defensive stats and rankings for each team. To merge these datasets, I looped through all files within the defensive datasets subfolder, and then added a column to identify which season the data corresponded too. Subsequent calculations were done to determine the relative pass defense rating of a team on a scale of 1-5. After merging, this dataset has 23 attributes and 480 tuples.
b.	Source: https://www.pro-football-reference.com/years/2013/opp.htm


```{r}
##Loop through all defensive datasets in data folder and merge
library(dplyr)
library(stringr)

# knitr::opts_chunk$set(echo = TRUE)
# knitr::opts_knit$set(root.dir = 'C:\\Users\\oshapira\\Desktop\\Analytics\\NFL_Quarterback_Performance\\Raw_Data\\defense_sportsref2000_2013') 


setwd('C:\\Users\\oshapira\\Desktop\\Analytics\\NFL_Quarterback_Performance\\Raw_Data\\defense_sportsref2000_2013')


file_list <- list.files()
file_list
# 
for (file in file_list){

  # if the merged dataset doesn't exist, create it
  if (!exists("defense_data")){
    defense_data <- read_excel(path = file, col_names = TRUE)
    defense_data$Year <- str_sub(toString(file), 9, 12)
  }

  # if the merged dataset does exist, append to it
  if (exists("defense_data")){
    temp_dataset <-read_excel(path = file, col_names = TRUE)
    temp_dataset$Year <- str_sub(toString(file), 9, 12)
    defense_data<-rbind(defense_data, temp_dataset)
    rm(temp_dataset)
  }

}

#Remove duplicate records created for Year 2000
defense_data <-defense_data %>% distinct()


kable(head(defense_data,3))
```


# **Data Scrubbing**

There was a significant amount of preprocessing that had to be done to prepare these datasets for data mining analysis. At a high level, this included filtering the datasets to rows/columns of interest, joining the datasets together, and discretizing numeric variables into categorical values. Preprocessing details for each dataset are described below:

```{r}
##import libraries
library(tidyr)
library(dplyr)
library(stats)
library(plotly)
```



## Weather dataset:
1.	Convert humidity variable from string to numeric
2.	Convert game date from string into date format
3.	Impute a value of 1 mph for all rows in the dataset that had a Wind mph value of 0. I felt there was a high likelihood the source data was inaccurate for these rows
4.	Remove teams from dataset that play in a dome. These teams/game locations are not subject to weather conditions, which was the primary data mining focus
5.	Convert team names to their respective acronyms (i.e. Buffalo Bills -> "BUF") so that this dataset could be joined to others that used team acronyms




```{r}
#convert humidity to a numeric variable
weather_raw$humidity_2 <- str_sub(weather_raw$humidity, 1, str_length(weather_raw$humidity)-1)
weather_raw$humidity_2 <- as.numeric(weather_raw$humidity_2)

##format date string into date type field
weather_raw$date_2<-strptime(weather_raw$date,format="%m/%d/%Y") #defining what is the original format of your date
weather_raw$date_2<-as.Date(weather_raw$date_2,format="%Y-%m-%d") #defining what is the desired format of your date

# replace all wind mph n/a values with 1
weather_raw$wind_mph[is.na(weather_raw$wind_mph)] <- 1
```


```{r}

###add column to specify teams that play in a dome, and therefore wouldn't be impacted by weather

weather_raw$dome <- "N"

weather_raw$dome[weather_raw$home_team == "Houston Oilers"] <-"Y"
weather_raw$dome[weather_raw$home_team == "New Orleans Saints"] <-"Y"
weather_raw$dome[weather_raw$home_team == "Detroit Lions"] <-"Y"
weather_raw$dome[weather_raw$home_team == "Houston Texans"] <-"Y"
weather_raw$dome[weather_raw$home_team == "Arizona Cardinals"] <-"Y"
weather_raw$dome[weather_raw$home_team == "Indianapolis Colts"] <-"Y"
weather_raw$dome[weather_raw$home_team == "Dallas Cowboys"] <-"Y"

###remove teams from dataset that play in a dome

weather_subset <- filter(weather_raw, weather_raw$dome == "N")

##add column that converts home team to an acronym so that it can later be joined to other datasets

weather_subset$home_team_2[weather_subset$home_team == "Arizona Cardinals"] <- "ARI"
weather_subset$home_team_2[weather_subset$home_team == "Atlanta Falcons"] <- "ATL"
weather_subset$home_team_2[weather_subset$home_team == "Baltimore Colts"] <- "BAL"
weather_subset$home_team_2[weather_subset$home_team == "Baltimore Ravens"] <- "BAL"
weather_subset$home_team_2[weather_subset$home_team == "Boston Patriots"] <- "BOS"
weather_subset$home_team_2[weather_subset$home_team == "Buffalo Bills"] <- "BUF"
weather_subset$home_team_2[weather_subset$home_team == "Carolina Panthers"] <- "CAR"
weather_subset$home_team_2[weather_subset$home_team == "Chicago Bears"] <- "CHI"
weather_subset$home_team_2[weather_subset$home_team == "Cincinnati Bengals"] <- "CIN"
weather_subset$home_team_2[weather_subset$home_team == "Cleveland Browns"] <- "CLE"
weather_subset$home_team_2[weather_subset$home_team == "Dallas Cowboys"] <- "DAL"
weather_subset$home_team_2[weather_subset$home_team == "Dallas Texans"] <- "DAL"
weather_subset$home_team_2[weather_subset$home_team == "Denver Broncos"] <- "DEN"
weather_subset$home_team_2[weather_subset$home_team == "Detroit Lions"] <- "DET"
weather_subset$home_team_2[weather_subset$home_team == "Green Bay Packers"] <- "GB"
weather_subset$home_team_2[weather_subset$home_team == "Houston Texans"] <- "HOU"
weather_subset$home_team_2[weather_subset$home_team == "Houston Oilers"] <- "HOU"
weather_subset$home_team_2[weather_subset$home_team == "Indianapolis Colts"] <- "IND"
weather_subset$home_team_2[weather_subset$home_team == "Jacksonville Jaguars"] <- "JAX"
weather_subset$home_team_2[weather_subset$home_team == "Kansas City Chiefs"] <- "KC"
weather_subset$home_team_2[weather_subset$home_team == "Miami Dolphins"] <- "MIA"
weather_subset$home_team_2[weather_subset$home_team == "Minnesota Vikings"] <- "MIN"
weather_subset$home_team_2[weather_subset$home_team == "New England Patriots"] <- "NE"
weather_subset$home_team_2[weather_subset$home_team == "New Orleans Saints"] <- "NO"
weather_subset$home_team_2[weather_subset$home_team == "New York Giants"] <- "NYG"
weather_subset$home_team_2[weather_subset$home_team == "New York Jets"] <- "NYJ"
weather_subset$home_team_2[weather_subset$home_team == "Oakland Raiders"] <- "OAK"
weather_subset$home_team_2[weather_subset$home_team == "Philadelphia Eagles"] <- "PHI"
weather_subset$home_team_2[weather_subset$home_team == "Phoenix Cardinals"] <- "PHO"
weather_subset$home_team_2[weather_subset$home_team == "Pittsburgh Steelers"] <- "PIT"
weather_subset$home_team_2[weather_subset$home_team == "Los Angeles Raiders"] <- "RAI"
weather_subset$home_team_2[weather_subset$home_team == "Los Angeles Rams"] <- "RAM"
weather_subset$home_team_2[weather_subset$home_team == "San Diego Chargers"] <- "SD"
weather_subset$home_team_2[weather_subset$home_team == "Seattle Seahawks"] <- "SEA"
weather_subset$home_team_2[weather_subset$home_team == "San Francisco 49ers"] <- "SF"
weather_subset$home_team_2[weather_subset$home_team == "St. Louis Cardinals"] <- "STL"
weather_subset$home_team_2[weather_subset$home_team == "St. Louis Rams"] <- "STL"
weather_subset$home_team_2[weather_subset$home_team == "Tampa Bay Buccaneers"] <- "TB"
weather_subset$home_team_2[weather_subset$home_team == "Tennessee Oilers"] <- "TEN"
weather_subset$home_team_2[weather_subset$home_team == "Tennessee Titans"] <- "TEN"
weather_subset$home_team_2[weather_subset$home_team == "Washington Redskins"] <- "WAS"
```



## Game Logs dataset:
1.	Filter on rows where QBs started. I wanted to exclude game log entries for QBs that likely didn't play many minutes and had very few pass attempts
2.	Convert game date from to proper date format
3.	Convert game week from integer to character
4.	Convert game completion percentage to a numeric variable
5.	Drop rows with NA values for game completion percentage 
6.	Filtered only on regular season games. Preseason games were excluded because QBs do not typically play their hardest in these games. Post-season games were excluded beacuse I thought data would be skewed to high-performing quarterbacks that may be less susceptible to weather. Post-season games also had game dates that spilled over into subsequent years (such as early January), which would have complicated my ability to join datasets based on season- year.
7.	Drop rows with NA values for passing yards


```{r}

#filter dataset on games where QB started
game_logs_played <- filter(game_logs_raw, game_logs_raw$Games.Started == 1)

#format date string into date type field
class(game_logs_played$Game.Date) #check original class of date field
game_logs_played$Game.Date_2 <- paste(game_logs_played$Game.Date, game_logs_played$Year, sep = "-")
game_logs_played$Game.Date_2 <-str_pad(game_logs_played$Game.Date_2, 11, pad = "0")
game_logs_played$Game.Date_2<-as.Date(game_logs_played$Game.Date_2, format = "%d-%b-%Y") #convert to date
class(game_logs_played$Game.Date_2)


##convert week column from integer to character
class(game_logs_played$Week)
game_logs_played$Week <- as.character(game_logs_played$Week)

##update team acronym for Jacksonville so that it matches the acronym in the weather dataset

game_logs_played$Opponent[game_logs_played$Opponent %in% "JAC"] <- "JAX"

```


## Career Stats dataset:

1. Flag players that normally play in domes for home games
2. Convert team names to their respective acronyms (i.e. Buffalo Bills -> "BUF") so that this dataset could be joined to others that used team acronyms
3. Drop rows with NA values for season completion percentage


```{r}
career$dome_player <- "N"

career$dome_player[career$Team == "Houston Oilers"] <-"Y"
career$dome_player[career$Team == "New Orleans Saints"] <-"Y"
career$dome_player[career$Team == "Detroit Lions"] <-"Y"
career$dome_player[career$Team == "Houston Texans"] <-"Y"
career$dome_player[career$Team == "Arizona Cardinals"] <-"Y"
career$dome_player[career$Team == "Indianapolis Colts"] <-"Y"
career$dome_player[career$Team == "Dallas Cowboys"] <-"Y"
```




```{r}

##convert team column to character
career$Team_2 <- as.character(career$Team)

##add a team acronym column so that it can be joined with other datasets

career$Team_2[career$Team_2 == "Arizona Cardinals"] <- "ARI"
career$Team_2[career$Team_2 == "Atlanta Falcons"] <- "ATL"
career$Team_2[career$Team_2 == "Baltimore Colts"] <- "BAL"
career$Team_2[career$Team_2 == "Baltimore Ravens"] <- "BAL"
career$Team_2[career$Team_2 == "Boston Patriots"] <- "BOS"
career$Team_2[career$Team_2 == "Buffalo Bills"] <- "BUF"
career$Team_2[career$Team_2 == "Carolina Panthers"] <- "CAR"
career$Team_2[career$Team_2 == "Chicago Bears"] <- "CHI"
career$Team_2[career$Team_2 == "Cincinnati Bengals"] <- "CIN"
career$Team_2[career$Team_2 == "Cleveland Browns"] <- "CLE"
career$Team_2[career$Team_2 == "Dallas Cowboys"] <- "DAL"
career$Team_2[career$Team_2 == "Dallas Texans"] <- "DAL"
career$Team_2[career$Team_2 == "Denver Broncos"] <- "DEN"
career$Team_2[career$Team_2 == "Detroit Lions"] <- "DET"
career$Team_2[career$Team_2 == "Green Bay Packers"] <- "GB"
career$Team_2[career$Team_2 == "Houston Texans"] <- "HOU"
career$Team_2[career$Team_2 == "Houston Oilers"] <- "HOU"
career$Team_2[career$Team_2 == "Indianapolis Colts"] <- "IND"
career$Team_2[career$Team_2 == "Jacksonville Jaguars"] <- "JAX"
career$Team_2[career$Team_2 == "Kansas City Chiefs"] <- "KC"
career$Team_2[career$Team_2 == "Miami Dolphins"] <- "MIA"
career$Team_2[career$Team_2 == "Minnesota Vikings"] <- "MIN"
career$Team_2[career$Team_2 == "New England Patriots"] <- "NE"
career$Team_2[career$Team_2 == "New Orleans Saints"] <- "NO"
career$Team_2[career$Team_2 == "New York Giants"] <- "NYG"
career$Team_2[career$Team_2 == "New York Jets"] <- "NYJ"
career$Team_2[career$Team_2 == "Oakland Raiders"] <- "OAK"
career$Team_2[career$Team_2 == "Philadelphia Eagles"] <- "PHI"
career$Team_2[career$Team_2 == "Phoenix Cardinals"] <- "PHO"
career$Team_2[career$Team_2 == "Pittsburgh Steelers"] <- "PIT"
career$Team_2[career$Team_2 == "Los Angeles Raiders"] <- "RAI"
career$Team_2[career$Team_2 == "Los Angeles Rams"] <- "RAM"
career$Team_2[career$Team_2 == "San Diego Chargers"] <- "SD"
career$Team_2[career$Team_2 == "Seattle Seahawks"] <- "SEA"
career$Team_2[career$Team_2 == "San Francisco 49ers"] <- "SF"
career$Team_2[career$Team_2 == "St. Louis Cardinals"] <- "STL"
career$Team_2[career$Team_2 == "St. Louis Rams"] <- "STL"
career$Team_2[career$Team_2 == "Tampa Bay Buccaneers"] <- "TB"
career$Team_2[career$Team_2 == "Tennessee Oilers"] <- "TEN"
career$Team_2[career$Team_2 == "Tennessee Titans"] <- "TEN"
career$Team_2[career$Team_2 == "Washington Redskins"] <- "WAS"
```





Defense dataset:
1.	Calculate mean, standard deviation, and z-score of each team's defensive strength (based on total passing yards allowed in the season). Metrics were calculated relative to all other teams' passing yards allowed within the same season. Since the NFL has gradually become more pass-heavy over time, I felt it was important to calculate z-scores of a team's defensive strength (passing yards allowed) as compared to other teams within the same season rather than across all seasons.
2. Use min-max normalization to convert a team's z-score value to a rating score between 1-5. A higher score indicates a stronger pass defense. Converting from z-score values that span between -3 to 3
3.	Drop columns that were irrelevant to my analysis (e.g. rushing yards allowed, pass completion %, etc.)
3.	Convert team names to their respective acronyms for purposes of joining with other datasets


```{r}
##Calculate average, standard deviation, and z-score of defensive strength

defense_data$avg <- sapply(seq_len(nrow(defense_data)), function(i) with(defense_data, mean(Yds[Year == Year[i]])))
defense_data$Standard_Deviation <- sapply(seq_len(nrow(defense_data)), function(i) with(defense_data, sd(Yds[Year == Year[i]])))
defense_data$Z_Score <-(defense_data$Yds - defense_data$avg)/defense_data$Standard_Deviation


##Convert to standard scale (1-5) across all season z-scores. So that it reflects if #1 team for one season is superior to #1 defense from prior season

def_z_min <- min(defense_data$Z_Score)
def_z_max <- max(defense_data$Z_Score)
def_z_range <- (def_z_max - def_z_min)
def_new_range <- 1 -5
defense_data$def_score <- (((defense_data$Z_Score - def_z_min)*def_new_range)/def_z_range)+5


##remove columns irrelevant for analysis
defense_data <- defense_data[ , -which(names(defense_data) %in% c("G", "Cmp", "Att", "Cmp%", "TD", "TD%", "Int", "Int%", "Y/A", "AY/A", "Y/C", "Y/G", "Rate", "Sk", "Yds__1", "NY/A", "ANY/A", "Sk%", "EXP"))]


##map team acronyms to team names so this can be joined to other datasets

defense_data$team_acr <- as.character(defense_data$Tm)

defense_data$team_acr[defense_data$team_acr == "Arizona Cardinals"] <- "ARI"
defense_data$team_acr[defense_data$team_acr == "Atlanta Falcons"] <- "ATL"
defense_data$team_acr[defense_data$team_acr == "Baltimore Colts"] <- "BAL"
defense_data$team_acr[defense_data$team_acr == "Baltimore Ravens"] <- "BAL"
defense_data$team_acr[defense_data$team_acr == "Boston Patriots"] <- "BOS"
defense_data$team_acr[defense_data$team_acr == "Buffalo Bills"] <- "BUF"
defense_data$team_acr[defense_data$team_acr == "Carolina Panthers"] <- "CAR"
defense_data$team_acr[defense_data$team_acr == "Chicago Bears"] <- "CHI"
defense_data$team_acr[defense_data$team_acr == "Cincinnati Bengals"] <- "CIN"
defense_data$team_acr[defense_data$team_acr == "Cleveland Browns"] <- "CLE"
defense_data$team_acr[defense_data$team_acr == "Dallas Cowboys"] <- "DAL"
defense_data$team_acr[defense_data$team_acr == "Dallas Texans"] <- "DAL"
defense_data$team_acr[defense_data$team_acr == "Denver Broncos"] <- "DEN"
defense_data$team_acr[defense_data$team_acr == "Detroit Lions"] <- "DET"
defense_data$team_acr[defense_data$team_acr == "Green Bay Packers"] <- "GB"
defense_data$team_acr[defense_data$team_acr == "Houston Texans"] <- "HOU"
defense_data$team_acr[defense_data$team_acr == "Houston Oilers"] <- "HOU"
defense_data$team_acr[defense_data$team_acr == "Indianapolis Colts"] <- "IND"
defense_data$team_acr[defense_data$team_acr == "Jacksonville Jaguars"] <- "JAX"
defense_data$team_acr[defense_data$team_acr == "Kansas City Chiefs"] <- "KC"
defense_data$team_acr[defense_data$team_acr == "Miami Dolphins"] <- "MIA"
defense_data$team_acr[defense_data$team_acr == "Minnesota Vikings"] <- "MIN"
defense_data$team_acr[defense_data$team_acr == "New England Patriots"] <- "NE"
defense_data$team_acr[defense_data$team_acr == "New Orleans Saints"] <- "NO"
defense_data$team_acr[defense_data$team_acr == "New York Giants"] <- "NYG"
defense_data$team_acr[defense_data$team_acr == "New York Jets"] <- "NYJ"
defense_data$team_acr[defense_data$team_acr == "Oakland Raiders"] <- "OAK"
defense_data$team_acr[defense_data$team_acr == "Philadelphia Eagles"] <- "PHI"
defense_data$team_acr[defense_data$team_acr == "Phoenix Cardinals"] <- "PHO"
defense_data$team_acr[defense_data$team_acr == "Pittsburgh Steelers"] <- "PIT"
defense_data$team_acr[defense_data$team_acr == "Los Angeles Raiders"] <- "RAI"
defense_data$team_acr[defense_data$team_acr == "Los Angeles Rams"] <- "RAM"
defense_data$team_acr[defense_data$team_acr == "San Diego Chargers"] <- "SD"
defense_data$team_acr[defense_data$team_acr == "Seattle Seahawks"] <- "SEA"
defense_data$team_acr[defense_data$team_acr == "San Francisco 49ers"] <- "SF"
defense_data$team_acr[defense_data$team_acr == "St. Louis Cardinals"] <- "STL"
defense_data$team_acr[defense_data$team_acr == "St. Louis Rams"] <- "STL"
defense_data$team_acr[defense_data$team_acr == "Tampa Bay Buccaneers"] <- "TB"
defense_data$team_acr[defense_data$team_acr == "Tennessee Oilers"] <- "TEN"
defense_data$team_acr[defense_data$team_acr == "Tennessee Titans"] <- "TEN"
defense_data$team_acr[defense_data$team_acr == "Washington Redskins"] <- "WAS"
```


## Joining datasets:
1.	Inner join Career Stats dataset with the Game Logs dataset based on matching season Year and QB name. Rename this "CareerGameLogs"
a.	Create additional column for game location using an if/else statement. If a game log was marked as "away" for a QB, then game location = Opponent. If it was a "home" game, then game location = the QB's team. 
i.	This added field was used to join with the weather dataset. Without this field, the weather dataset would have only joined with the merged GameLogs/CareerStats dataset on games where the QB was playing a home game. 	
b.	Create a "Yards deviation" column that showed how a QB's passing yards in a game deviate from their season average (game log passing yards - season average passing yards). This was the primary "performance" metric which was later discretized to be the class variable we tried to predict.
2.	Join the merged "CareerGameLogs" with the weather dataset, based on matching game date and game location. Rename this "weather_gamelogs_inner".
3.	Join the merged "weather_gamelogs_inner" with the Defense dataset by joining on matching season year and team acronym.


```{r}

################################
# join the career stats with the game logs
################################

#Note: The join actually added some rows. This is due to some QBs (<100 rows) that were traded mid-season, and have multiple rows for different teams in the same season
careerGameLogs <- merge(x=game_logs_played, y=career, by.x = c("Year","Name"), by.y = c("Year","Name"), all.x = TRUE)


##drop career stat columns that were irrelevent for analysis
careerGameLogs_2 <- subset(careerGameLogs, select=-c(Games.Played.y,Passes.Attempted.y,Passes.Completed.y,Passing.Yards.y,TD.Passes.y,Percentage.of.TDs.per.Attempts,Ints.y,Sacks.y,Sacked.Yards.Lost.y))


##Add Game_location field to the gamelog column, which will be used for joining to the weather dataset
careerGameLogs_2$game_location <- careerGameLogs_2$Team

#If a QB's game log is marked as an away game, then location = opponent team. If game log is home game for QB, location = QB team
careerGameLogs_2$game_location <- ifelse(careerGameLogs_2$Home.or.Away == "Away",as.character(careerGameLogs_2$Opponent),careerGameLogs_2$Team_2)

################################
##Merge weather and gamelogs dataset
################################

##inner join
weather_gamelogs_inner <- merge(x=weather_subset, y=careerGameLogs_2, by.x = c("date_2", "home_team_2"), by.y = c("Game.Date_2", "game_location") , by = NULL)


########clean up inner join file######################


##remove irrelevant columns
weather_gamelogs_inner_2 <- subset(weather_gamelogs_inner, select=-c(id, date, wind_chill, humidity, weather, dome, Player.Id.x, Position.x, Game.Date, Games.Played.x, Games.Started, Player.Id.y, Position.y))

#relabel column names so that they more clearly represent what they are
colnames(weather_gamelogs_inner_2) <- c("game_date", "game_location_acr", "home_team", "home_score", "away_team", "away_score", "temperature", "wind_mph", "humidity_pct", "year", "qb_name", "season", "week", "home_or_away", "opponent_acr", "outcome", "score", "game_passes_completed", "game_passes_attempted", "game_completion_pct", "game_passing_yards", "game_yards_per_attempt", "game_td_passes", "game_interceptions", "game_sacks", "game_sacked_yards_lost", "game_passer_rating", "game_rushing_attempts", "game_rushing_yards", "game_yards_per_carry", "game_rushing_td", "game_fumbles", "game_fumbles_lost", "qb_team", "season_completion_pct", "season_attempts_per_game", "season__yards_per_attmpt", "season_yards_per_game", "season_interecption_rate", "season_longest_pass", "season_passes_longer_20yards", "season_passes_longer_40yards", "season_passer_rating", "dome_player", "qb_team_acr")

#re-order colums
weather_gamelogs_inner_3 <- weather_gamelogs_inner_2[, c("qb_name", "qb_team", "qb_team_acr", "game_date", "week", "season", "opponent_acr", "home_or_away", "away_team", "game_location_acr", "home_team", "home_score", "away_score", "outcome", "score", "temperature", "wind_mph", "humidity_pct", "year", "game_passes_completed", "game_passes_attempted", "game_completion_pct", "game_passing_yards", "game_yards_per_attempt", "game_td_passes", "game_interceptions", "game_sacks", "game_sacked_yards_lost", "game_passer_rating", "game_rushing_attempts", "game_rushing_yards", "game_yards_per_carry", "game_rushing_td", "game_fumbles", "game_fumbles_lost", "season_completion_pct", "season_attempts_per_game", "season__yards_per_attmpt", "season_yards_per_game", "season_interecption_rate", "season_longest_pass", "season_passes_longer_20yards", "season_passes_longer_40yards", "season_passer_rating", "dome_player")]


#Add column that calculates deviation of a QB's passing yards per game compared to their yards/per game average within that same season

weather_gamelogs_inner_3$yards_dev <- as.numeric(as.character(weather_gamelogs_inner_3$game_passing_yards)) - as.numeric(as.character(weather_gamelogs_inner_3$season_yards_per_game))





###merge defense dataset with other merged datasets
weather_gamelogs_def_inner <- merge(x=weather_gamelogs_inner_3, y=defense_data, by.x = c("year", "opponent_acr"), by.y = c("Year", "team_acr") , by = NULL)


# ##normalizing passing yards per game based on defensive strength
# weather_gamelogs_def_inner$game_passing_yards_norm<-as.numeric(as.character(weather_gamelogs_def_inner$game_passing_yards))/as.numeric(as.character(weather_gamelogs_def_inner$Min_max_norm))

# #derive yards deviation from season average, with normalized game passing yards
# weather_gamelogs_def_inner$yards_dev_norm <- as.numeric(as.character(weather_gamelogs_def_inner$game_passing_yards_norm)) - as.numeric(as.character(weather_gamelogs_def_inner$season_yards_per_game))
# 
# 
# 
# ##drop rows with NA humidity
# weather_gamelogs_def_inner <- weather_gamelogs_def_inner[!is.na(weather_gamelogs_def_inner$humidity_pct),]
# 
# ##drop rows with NA yards deviation
# weather_gamelogs_def_inner <- weather_gamelogs_def_inner[!is.na(weather_gamelogs_def_inner$yards_dev),]



###filter dataset on only regular season games. Exclude playoffs and preseason
weather_gamelogs_def_inner <- filter(weather_gamelogs_def_inner, weather_gamelogs_def_inner$season == "Regular Season")



```



```{r}
# setwd('C:\\Users\\oshapira\\Desktop\\Analytics\\NFL_Quarterback_Performance\\scrubbed_data')
write.csv(weather_gamelogs_def_inner, 'C:\\Users\\oshapira\\Desktop\\Analytics\\NFL_Quarterback_Performance\\scrubbed_data\\scrubbed_data.csv')

```

